---
title: "Annotating Treatments Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Annotating Treatments Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# THIS VIGNETTE IS A WORK IN PROGRESS

## This is an example Pipeline to annotate treatments from the CTRP dataset. 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(AnnotationGx)
library(data.table)
options(mc.cores = parallel::detectCores())
```


### Example pipeline to annotate a dataseGSM1897356t
#TODO::incomplete

We will be working with some data from the GDSC and Cell Model Passports datasets 
for this vignette. The GDSC dataset contains information about the cell lines 
in the Genomics of Drug Sensitivity in Cancer (GDSC) project. The Cell Model
Passports dataset contains information about all the models in the the database.


The GDSC sampleMetadata contains two columns, `GDSC.Sample_Name` and `GDSC.COSMIC_ID`.
``` {r setup_data}
data(GDSC_sampleMetadata)
data(gCSI_sampleMetadata)
data(CTRP_sampleMetadata)
data(CCLE_sampleMetadata)

cell_lines_all <- data.table::fread("/home/bioinf/bhklab/jermiah/Bioconductor/AnnotationGx/inst/extdata/cell_annotation_all.csv")

```

``` {r view_GDSC_sampleMetadata}

head(GDSC_sampleMetadata)
name <- GDSC_sampleMetadata$GDSC.Sample_Name
keep_duplicates <- F
result <- mapCell2Accession(
  name
  )
GDSC_result <- result
setnames(GDSC_result, "query", "GDSC.sampleid")
GDSC_result

data.table::fwrite(GDSC_result, "GDSC_result.csv")
```



``` {r view_gCSI_sampleMetadata}

head(gCSI_sampleMetadata)

name <- gCSI_sampleMetadata[["Characteristics.cell.line."]]
keep_duplicates <- F
result <- mapCell2Accession(
  name
  )
gCSI_result <- result
setnames(gCSI_result, "query", "gCSI.sampleid")

print("Missing samples in gCSI")
name <- gCSI_result[is.na(accession),unique(gCSI.sampleid)]
result <- mapCell2Accession(name, fuzzy = T)
setnames(result, "query", "gCSI.sampleid")
gCSI_result <- data.table::rbindlist(
  list(
    unique(gCSI_result[!is.na(accession)]),
    result
  )
) |> unique()

gCSI_result
data.table::fwrite(gCSI_result, "gCSI_result.csv")
```


``` {r combine_gdsc_gcsi}


merged <- unique(merge(gCSI_result, GDSC_result, c("cellLineName", "accession"), all = T))
merged <- merged[order(cellLineName)]

print("samples in gCSI that are NOT in GDSC")
print(merged[is.na(GDSC.sampleid), unique(gCSI.sampleid)])

print("samples in GDSC NOT in gCSI")
print(merged[is.na(gCSI.sampleid), unique(GDSC.sampleid)])

print("samples in both GDSC and gCSI")
print(merged[!is.na(GDSC.sampleid) & !is.na(gCSI.sampleid),])



subset_all <- cell_lines_all[, .(gCSI.cellid, GDSC_rnaseq.cellid,GDSC1000.cellid, Cellosaurus.Accession.id)]
# remove any rows with all ""
subset_all <- subset_all[gCSI.cellid != "" | GDSC_rnaseq.cellid != "" | GDSC1000.cellid != ""]
names(subset_all) <- paste0("OLD_", names(subset_all))
merged <- merge(merged, subset_all, by.x = "accession", by.y = "OLD_Cellosaurus.Accession.id", all.x = T)
merged <- merged[order(cellLineName)]
# new column order
cols <- c("cellLineName", "accession", "gCSI.sampleid", "OLD_gCSI.cellid", "GDSC.sampleid", "OLD_GDSC_rnaseq.cellid", "OLD_GDSC1000.cellid")
merged <- merged[, ..cols]
data.table::fwrite(merged, "merged.csv")

```

``` {r ccle}
head(CCLE_sampleMetadata)
CCLE_result <- mapCell2Accession(CCLE_sampleMetadata$CCLE_ID_parsed)
CCLE_result

CCLE_failed <- mapCell2Accession(CCLE_result[is.na(accession),query], fuzzy = T, parsed = F)
CCLE_result <- data.table::rbindlist(
  list(
    unique(CCLE_result[!is.na(accession)]),
    unique(CCLE_failed[!is.na(accession)])
  )
) |> unique()

CCLE_failed_again_names <- CCLE_failed[is.na(accession),query]

depmapids <- na.omit(CCLE_sampleMetadata[CCLE_ID_parsed %in% CCLE_failed_again_names, depMapID])
failed_depmaps <- mapCell2Accession(depmapids, from = "dr")
failed_depmaps$query <- CCLE_sampleMetadata[match(depmapids, CCLE_sampleMetadata$depMapID), CCLE_ID_parsed]
CCLE_result <- data.table::rbindlist(
  list(
    unique(CCLE_result[!is.na(accession)]),
    unique(failed_depmaps[!is.na(accession)])
  )
) |> unique()
CCLE_result[order(cellLineName)]
missing <- setdiff(CCLE_sampleMetadata$CCLE_ID_parsed, CCLE_result$query)
message("The following is still missing: ", missing)

data.table::fwrite(CCLE_result, "CCLE_result.csv")

depmapids_mapped <- mapCell2Accession(CCLE_sampleMetadata[!is.na(depMapID), depMapID], from = "dr")
depmapids_mapped[order(cellLineName)]
data.table::fwrite(depmapids_mapped, "CCLE_depmapids_mapped.csv")

```


``` {r ctrp}
head(CTRP_sampleMetadata)

CTRP_result <- mapCell2Accession(CTRP_sampleMetadata$ccl_name)

CTRP_failed <- mapCell2Accession(CTRP_result[is.na(accession),query], fuzzy = T, parsed = F)

CTRP_failed_again_names <- CTRP_failed[is.na(accession),query]

message("The following is still missing: ", paste0(CTRP_failed_again_names, collapse = ", "))

data.table::fwrite(CTRP_result, "CTRP_result.csv")

```


``` {r combine_all}
# read in the 4 dataset new csvs
GDSC_result <- data.table::fread("TRASH/GDSC_result.csv")
gCSI_result <- data.table::fread("TRASH/gCSI_result.csv")
CCLE_result <- data.table::fread("TRASH/CCLE_result.csv")
CTRP_result <- data.table::fread("TRASH/CTRP_result.csv")




```