---
title: "Querying Cellosaurus"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Querying Cellosaurus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
Cellosaurus is a comprehensive knowledge resource dedicated to cell lines,
providing a wealth of information about various types of cells used in
biomedical research. It serves as a centralized repository that offers
detailed data on cell lines, including their origins, characteristics,
authentication methods, references, and more. Please view the Cellosaurus
website at https://web.expasy.org/cellosaurus/ for more information and
a detailed description can be found at https://www.cellosaurus.org/description.html.

The `AnnotationGx` package provides a wrapper around the Cellosaurus API to 
map cell line identifiers to the Cellosaurus database fields. 

## Setup

```{r setup}
library(AnnotationGx)
library(data.table)

# set options to warn to quiet info logs
options("log_level" = "WARN")
```


## Mapping from Cell Line name to Accession id
The main function that is provided by the package is `mapCell2Accession`. This function
takes in a vector of cell line identifiers and returns a `data.table`. 

By default, the function will try to map using the common identifiers and synonyms (`from = "idsy"`) and 
will return the the Standardized Identifier as `cellLineName` and the Cellosaurus Accession ID `accession`. 
The function also returns an additional column `query` which can be used to identify the original query if needed. 

Let's see how we can use this function to map the "HeLa" and "A549" cell line names
to the Cellosaurus database. 
```{r map heLa each}
mapCell2Accession("hela")
```

```{r map A549 each}
mapCell2Accession("A549")
```

Functionality for mapping multiple cell lines is also supported.
``` {r map BT474 each}
mapCell2Accession(c("A549", "THIS SHOULD FAIL", "BT474"))
```


By default, the function will parse the API responses to return the most common mapping. 
To return all possible mappings, set `parsed = FALSE`.

```{R parsed}
# parsed
mapCell2Accession(c("A549", "hela", "BT474"), parsed = TRUE)

# no parsing
mapCell2Accession(c("A549", "hela", "BT474"), parsed = FALSE)
```


### Misspellings and synonyms
The backend of the function also tries to map any misspellings or synonyms of the cell line names.
```{R mispellings}

samples <- c("SK23", "SJCRH30")
mapCell2Accession(samples)

```


If some cell lines still cannot be found, there is an additional parameter for fuzzy searching 
```{R fuzzy}

# No fuzzy 
mapCell2Accession("DOR 13")

# Fuzzy
mapCell2Accession("DOR 13", fuzzy =T)

```



## Annotating Cellosaurus Accessions

Once accession IDs are obtained and the mappings are satisfactory, they can then be mapped to other fields in the Cellosaurus database.
A list of available fields can be found using `cellosaurus_fields()`

```{R fields}
cellosaurus_fields()
```


The `annotateCellAccession()` function can be used to map the accession IDs to the desired fields.
By default the function will try to map to `"id", "ac", "hi", "sy", "ca", "sx", "ag", "di", "derived-from-site", "misspelling", "dt"`

```{R annotate}

# Annotate the A549 cell line
mappedAccessions <- mapCell2Accession("A549")

annotateCellAccession(accessions = mappedAccessions$accession)

```
