---
title: "Querying Cellosaurus"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Querying Cellosaurus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
Cellosaurus is a comprehensive knowledge resource dedicated to cell lines,
providing a wealth of information about various types of cells used in
biomedical research. It serves as a centralized repository that offers
detailed data on cell lines, including their origins, characteristics,
authentication methods, references, and more. Please view the Cellosaurus
website at https://web.expasy.org/cellosaurus/ for more information and
a detailed description can be found at https://www.cellosaurus.org/description.html.

The `AnnotationGx` package provides a wrapper around the Cellosaurus API to 
map cell line identifiers to the Cellosaurus database fields. 

### Setup

```{r setup}
library(AnnotationGx)
library(data.table)
```


``` {r setBPPARAM}
BPPARAM = BiocParallel::MulticoreParam(progressbar=T)
```
### Mapping from Cell Line name to Accession id
The main function that is provided by the package is `mapCell2Accession`. This function
takes in a vector of cell line identifiers and returns a `data.table`. 

```
Usage and default values:
     mapCell2Accession(
       ids,
       numResults = 1000,
       from = "id",
       to = c("id", "ac"),
       prioritizeParent = FALSE,
       orderby = "ac",
       query_only = FALSE,
       raw = FALSE,
       BPPARAM = BiocParallel::SerialParam(),
       ...
     )
```

Let's see how we can use this function to map the "HeLa" and "A549" cell line names
to the Cellosaurus database. 
By default, the function will try to map using the common identifier (`from ="id"`) name and 
will return the standard identifier `id` and the Cellosaurus accession ID `ac`. The 
function also returns two additional columns `query` and `query:<from>` which
can be used to identify the original query and the original identifier type if needed. 
``` {r map heLa each}
(result <- mapCell2Accession("HeLa", keep_duplicates = T, raw = T))
result$HeLa -> resp

(result <- mapCell2Accession("HeLa", keep_duplicates = T, raw = F))

```


``` {r map A549 each}
mapCell2Accession("A549")
```

Another example using the BT474 cell line. 

``` {r map BT474 each}
# our query:
mapCell2Accession("BT474")

mapCell2Accession(c("A549", "OOPS", "BT474"))

```
