---
title: "Querying ChEMBL Database"
author: 
  - name: Jermiah Joseph, Shahzada Muhammad Shameel Farooq, and Christopher Eeles
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('AnnotationGx')`"
vignette: >
  %\VignetteIndexEntry{Querying ChEMBL Database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL ## Related to https://stat.ethz.ch/pipermail/bioc-devel/2020-April/016656.html
)
```


# Introduction to ChEMBL API

**WARNING: This vignette is a work in progress. If you have questions or would
like to see more features, please open an issue at 
[bhklab/AnnotationGx](https://github.com/bhklab/AnnotationGx)**

The ChEMBL database contains information on bioactive drug-like small molecules.
The information includes 2-D structures, calculated properties; logP, Molecular
Weight, Lipinski Parameters, and abstracted bioactivities; binding
constants and ADMET data. The data is curated from primary scientific literature.
The ChEMBL API allows for the data to be made available for retrieval in a
programmatic fashion. We can use the API to query CHEMBL ID of a compound, retrieve
all molecule mechanisms of action, query compound_record resource and molecule
resource from the ChEMBL database.


## Setup

```{r setup agx}
library(AnnotationGx)
```



## Retrieve molecule mechanisms of action from ChEMBL {#chembl-mechanisms}

Given a ChEMBL ID, we can retrieve the molecule mechanisms of action from the
ChEMBL database using the `getChemblMechanism()` function. 

**
NOTE: This is a specialized function that queries the API for the *mechanism* resource only.
To query other resources, please see the [Custom Queries](#custom-queries) section.
**



``` {r run one query}
mechs <- getChemblMechanism("CHEMBL1413")
mechs
```

In the above example, multiple mechanisms of action are returned. 


## Custom Queries {#custom-queries}

The ChEMBL API allows for a wide range of queries. We have specialized one function,
but are open to incorporating more. Please open an issue at [bhklab/AnnotationGx](https://github.com/bhklab/AnnotationGx)
with an idea of a specialized function that meets a use case. 

A query to the API follows the following format:
```
https://www.ebi.ac.uk/chembl/api/data/[resource]?[field]__[filter_type]=[value]&format=[format]
```
More information can be found at the [API Documentation](https://chembl.gitbook.io/chembl-interface-documentation/web-services/chembl-data-web-**services******)

In summary, the requirements for a query are:

1. The `resource` to be queried
2. The reource `field` to be queried
3. The `filter_type` to be used
4. The `value` to be used for the filter
5. (optional) The `format` of the returned data (default is JSON)

For example, the query for the example in [the above section](#chembl-mechanisms) would be:
"https://www.ebi.ac.uk/chembl/api/data/mechanism?molecule_chembl_id__in=CHEMBL1413&format=json"
where:

- `resource` is "mechanism"
- `field` is "molecule_chembl_id"
- `filter_type` is "in"
- `value` is "CHEMBL1413"
- `format` is "json"

These parameters can be used in the `queryChemblAPI(resource, field, filter_type, value, format = "json")`
function to query the ChEMBL API.

**NOTE: unlike the `getChemblMechanism()` function which returns a `data.table`, the `queryChemblAPI()` function 
returns the raw data unformatted**

``` {r run custom query}
queryChemblAPI("mechanism", "molecule_chembl_id", "in", "CHEMBL1413")
```

The `getChemblResources()` function returns a list of possible 
resources that can be queried:

``` {r query resources}
getChemblResources() 
```


The `getChemblResourceFields(resource)` function returns a list of possible
fields that can be queried for a given resource:
``` {r query resource fields}
getChemblResourceFields("mechanism")

```


The `getChemblFilterTypes()` function returns a list of possible filter types.

``` {r query filter types}
getChemblFilterTypes()
```
