---
title: "Introduction to AnnotationGx"
author:
  - name: Jermiah Joseph
    affiliation:
    - BHKLab
output:
  BiocStyle::html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to AnnotationGx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup-options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  crop = NULL
)
```

# Introduction

`AnnotationGx` provides utilities for mapping pharmacogenomic sample and
treatment identifiers to public annotation resources. The package focuses on
the practical annotation tasks that recur when harmonizing drug response
studies: mapping cell line names to Cellosaurus accessions, mapping treatment
names to PubChem compound identifiers, collecting selected ChEMBL/UniChem
annotations, and querying cancer type vocabularies such as OncoTree.

The package belongs in Bioconductor because it supports reproducible
annotation of pharmacogenomic data objects and complements packages such as
`PharmacoGx`, which store and analyze pharmacogenomic experiments. `AnnotationGx`
is intended to sit near the data preparation boundary: it helps normalize
external identifiers before downstream analyses are performed.

# Installation

Install the package from Bioconductor with:

```{r install-bioc, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
BiocManager::install("AnnotationGx")
```

During development, install the current source checkout from a terminal with
`R CMD INSTALL .`.

# Scope and Related Packages

`AnnotationGx` is narrower than general annotation frameworks such as
`AnnotationDbi` and narrower than assay containers such as `PharmacoGx`. It
wraps selected web services that are commonly needed for pharmacogenomic sample
and treatment harmonization. It does not replace those packages.

The small `BioMartClient` helper in this package is currently used for a
narrow HGNC BioMart REST workflow and shares the package's caching and request
handling helpers. If BioMart usage expands beyond this focused workflow, the
implementation should be reconsidered against the Bioconductor `biomaRt`
package.

# Core Functionality

Load the package and inspect the small example metadata tables:

```{r load-package, message = FALSE}
library("AnnotationGx")
data(GDSC_sampleMetadata)
data(GDSC_treatmentMetadata)

head(GDSC_sampleMetadata)
head(GDSC_treatmentMetadata)
```

Build Cellosaurus queries for cell line identifiers:

```{r cellosaurus-query}
mapCell2Accession(c("A549", "BT474"), query_only = TRUE)
```

Run the query when an internet connection is available:

```{r cellosaurus-online}
mapCell2Accession(c("A549", "BT474"))
```

Build PubChem PUG-REST queries for treatment names:

```{r pubchem-query}
mapCompound2CID(c("erlotinib", "paclitaxel"), query_only = TRUE)
```

Run the PubChem query when online examples are enabled:

```{r pubchem-online}
mapCompound2CID(c("erlotinib", "paclitaxel"), first = TRUE)
```

Map a PubChem CID to selected computed properties:

```{r pubchem-properties}
mapCID2Properties(
  ids = c(176870, 36314),
  properties = c("Title", "MolecularFormula", "InChIKey")
)
```

Inspect available Cellosaurus fields used by the cell line annotation helpers:

```{r cellosaurus-fields}
head(cellosaurus_fields())
```

# Data Provenance

The example data objects are derived from public pharmacogenomic resources and
are documented individually in their help pages. The source package also
includes `inst/extdata/LICENSES.md`, which records provenance, source links,
and redistribution notes for bundled data.

```{r data-license-path}
system.file("extdata", "LICENSES.md", package = "AnnotationGx")
```

# Citing AnnotationGx

```{r citation}
citation("AnnotationGx")
```

# Session Info

```{r session-info}
sessionInfo()
```
