---
title: "Introduction to AnnotationGx"
author: 
  - name: Jermiah Joseph
    affiliation:
    - BHKLab 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Introduction to AnnotationGx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup-options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  crop = NULL ## Related to https://stat.ethz.ch/pipermail/bioc-devel/2020-April/016656.html
)
```

# Basics

## Install `AnnotationGx`

`R` is an open-source statistical environment which can be easily modified to 
enhance its functionality via packages. `AnnotationGx` is a `R` package hosted
on Bioconductor and can be installed using `BiocManager`:

```{r install}
BiocManager::install("AnnotationGx")
```

## Citing `AnnotationGx`

We hope that `AnnotationGx` will be useful for your research. Please use the 
following information to cite the package and the overall approach. Thank you!

```{r citation}
## Citation info
citation("AnnotationGx")
```

# Quick start to using `AnnotationGx`

```{r load-package, message=FALSE}
library("AnnotationGx")
```

```{r session-info}
sessionInfo()
```
