---
title: "Querying OncoTree"
output:
  BiocStyle::html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Querying OncoTree}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup-options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
online_examples <- nzchar(Sys.getenv("ANNOTATIONGX_RUN_ONLINE_EXAMPLES"))
```

# Introduction

OncoTree is a standardized classification system used in cancer research 
and clinical practice to categorize different types of cancer based on 
their tissue of origin, molecular characteristics, and other relevant factors. 
Developed by Memorial Sloan Kettering Cancer Center (MSK), 
OncoTree provides a hierarchical framework that organizes cancer types into
a structured tree-like diagram.

- provides a standardized classification system for categorizing different
  types of cancer based on their tissue of origin, molecular characteristics,
  and other relevant factors.
- provides a hierarchical framework that organizes cancer types into a 
  structured tree-like diagram.
- useful for ensuring consistency in how cancer types are classified and 
  reported across different studies and clinical settings.

## Licensing

OncoTree is licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/). Source: https://github.com/cBioPortal/oncotree


# Setup

```{r setup}
library(AnnotationGx)
```

# Querying OncoTree

AnnotationGx provides a set of functions for querying OncoTree to retrieve
three types of information:
- OncoTree release versions
- Main Cancer types
- Subtypes of a specific cancer type and their relationships

## OncoTree release versions

The `getOncotreeVersions` function retrieves the available OncoTree release.
```{r oncotree-versions, eval = online_examples}
getOncotreeVersions()
```

## Main Cancer types

The `getMainCancerTypes` function retrieves the main cancer types in OncoTree.
```{r main-cancer-types, eval = online_examples}
getOncotreeMainTypes()
```

## Subtypes of a specific cancer type

The `getCancerSubtypes` function retrieves the subtypes of a specific cancer 
type.
```{r cancer-subtypes, eval = online_examples}
getOncotreeTumorTypes()
```

```{r session-info}
sessionInfo()
```
