---
title: "Querying OncoTree"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Querying OncoTree}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
OncoTree is a standardized classification system used in cancer research 
and clinical practice to categorize different types of cancer based on 
their tissue of origin, molecular characteristics, and other relevant factors. 
Developed by the National Cancer Institute (NCI) within the United States, 
OncoTree provides a hierarchical framework that organizes cancer types into
a structured tree-like diagram.

- provides a standardized classification system for categorizing different
  types of cancer based on their tissue of origin, molecular characteristics,
  and other relevant factors.
- provides a hierarchical framework that organizes cancer types into a 
  structured tree-like diagram.
- useful for ensuring consistency in how cancer types are classified and 
  reported across different studies and clinical settings.


## Setup
```{r setup}
library(AnnotationGx)
```

## Querying OncoTree

AnnotationGx provides a set of functions for querying OncoTree to retrieve
three types of information:
1) OncoTree release versions
2) Main Cancer types
3) Subtypes of a specific cancer type and their relationships

### OncoTree release versions
The `getOncotreeVersions` function retrieves the available OncoTree release.
```{r getOncotreeVersions}
getOncotreeVersions()
```

### Main Cancer types
The `getMainCancerTypes` function retrieves the main cancer types in OncoTree.
```{r getMainCancerTypes}
getOncotreeMainTypes()
```

### Subtypes of a specific cancer type
The `getCancerSubtypes` function retrieves the subtypes of a specific cancer type.
```{r getCancerSubtypes}
getOncotreeTumorTypes("Lung")
```


