---
title: "Querying PubChem with AnnotationGx"
author: "Christopher Eeles"
package: AnnotationGx
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

# Introduction to PubChem APIs

PubChem provides two publicly available RESTful APIs for programatically
accessing data in their databases. The first of these, the PubChem PUG REST API,
is designed to make specific queries based on some input identifier and return
data which PubChem has labelled or computed internally [1]. This API can be used 
to do things like get standardized PubChem identifiers CID or SID from 
drug names, external database identifiers, etc. The second is the PubChem PUG 
VIEW API [2]. This interface allows access to a variety of data which are 
aggregated by PubChem to display along with their own records, but are not 
curated by PubChem itself.

## PubChem PUG REST API

This package inlcudes functions for making arbitrary queries to the PubChem
REST API. The most fundamental of these is the `getRequestPubChem` function, 
which uses its parameters to construct a single GET request to the PUG REST API, 
returning an `httr::response` object. The documentation for this function
also includes the most detail about the structure of the PUG REST API, and
developers wishing to write custom queries for this package should review
the documentation with `?getRequestPubChem`.

For convenience, a wrapper function, `queryRequestPubChem`, is available which 
automatically converts the results to a `data.frame` using `jsonlite::fromJSON`. 
This function only works when `output` parameter is set to `JSON`. 

The major workhorse for PubChem queries in this package is the `queryPubChem` 
function. This function is able to batch queries such that multiple identifiers 
are mapped per GET request. It also implementes parallelized queries as well as
query throttling to ensure users do not exceed the 5 query per second, 400
query per minute limits imposed by PubChem. Both 
`queryRequestPubChem` use the `...` parameter to pass
arguments through to `getRequestPubChem`. As such, users should review the
`getRequestPubChem` documentation prior to using either query functions.

Two additional parameters are added in `queryPubChem`. The `batch` query 
tells the function whether to use multiple identifiers per GET request. This
parameter should be set to `TRUE` (the default) unless you are retrying failed 
queries. This is because batch queries fail completely if ANY of their
their ids fail to map. The `raw` parameter is included for debugging purposes,
This can be used to diagnose HTTP errors using the `headers(response)` function.

Use of `queryPubChem` is mainly intended for developer use, and beginners
with this package should focus on using the more specialized functions described
below. It is worth noting, however, that all specialized functions were 
constructed using `queryPubChem`.

The PUG REST API is divided into multiple domains, each representing a distinct
entity in the PubChem Schema. Details of these domains can be found in 
`?getRequestPubChem`. The three most important domains are discussed below.

### Querying the Compound Domain

The compound domain allows users to make queries on pharmacologically relevant
entities such as drugs. The main PubChem id for compounds is Compound IDentifier 
(CID). Each CID may map to one or more Substance IDentifiers (SID). In this
context, SID is a less stable identifier, possibly only contributed
by a single source and not yet standardized. CIDs on the otherhand, have stable
PubChem entries and normalized names which are generally supported accross 
multiple data sources. As such, newly discovered or rarely used chemicals
may have an SID but not a CID.

For most users, the compound domain will provide the bulk of the information
they need about their drugs of interest. The `getPubChemCompound` function
has been included to allow users to easily query this domain.

#### Common Use Cases

1. Get CID from Drug Names

```{r eval=FALSE}
library(AnnotationGx)
drugs <- c("Erlotinib", "Acadesine", "Camptothecin", "Vincaleukoblastine", 
"Cisplatin")
namesToCID <- getPubChemCompound(drugs, from='name', to='cids')
```


2. Get SID from Drug Names
```{r eval=FALSE}
namesToSID <- getPubChemCompound(drugs, from='name', to='sids')
```


3. Get Synonyms from Drug Names
```{r eval=FALSE}
namesToSynonyms <- getPubChemCompound(drugs, from='name', to='synonyms')
```


4. Get Compound Properties from CIDs
```{r eval=FALSE}
cids <- namesToCID$cids
propertiesFromCID <- getPubChemCompound(cids, 
    properties=c('Title', 'InChIKey', 'CanonicalSMILES'))
```

A short-list of properties are available in the function documentation, see
`?getPubChemCompound`. A link in the DETAILS section will take you to a full
list of available compound properties.

### Querying the Substance Domain

The substance domain allows users to query a broader range of chemicals than
the compounds domain. This can be useful if mappings from name or CID failed
to get the information you are interested in.

### Querying the Assay Domain

A functions for making specific queries to the assay domain on the PUG REST API
has not been implemented yet. A user could implement their own using the
`queryPubChem` functions and knowledge of the PubChem API.

### Specialized functions

To make the package more user friendly, we have included a number of convenience
functions which further limit the scope of queries, but require less knowledge
of the underlying concepts of the PubChem API. We hope that this will allow
users with no knowledge or interest in the PubChem API to retrieve useful
information nonetheless.

1. Mapping from NSC Number to CIDs

```{r eval=FALSE}
NSC_numbers <- c(1026L, 1027L, 1042L, 1059L, 1063L, 1139L)
NSCtoCID <- getPubChemFromNSC(NSC_numbers, to='cids')
```

## PubChem PUG VIEW API

The PUG View API is primarily designed to complete summary reports on individual
PubChem records.

### Querying Annotations

The PubChem PUG View Annotations API provides summaries of third-party 
annotations accross all PubChem records. These annotations are linked To
PubChem compounds via CID and/or SID. This API does not allow specific
input queries, instead returning the entire record for the user to parse
and extract relevant information.

1. Getting All ATC Codes

```{r eval=FALSE}
CIDtoATC <- getPubChemAnnotations('ATC Codes')
```

2. Getting All NSC Numbers

```{r eval=FALSE}
CIDtoNSC <- getPubChemAnnotations('NSC Number')
```

3. Getting All CTD Codes

```{r }
CIDtoNSC <- getPubChemAnnotations('NSC Number')
```


# References
1. PUG REST. PubChem Docs [website]. Retrieved from https://pubchemdocs.ncbi.nlm.nih.gov/pug-rest.
2. Kim S, Thiessen PA, Cheng T, Yu B, Bolton EE. An update on PUG-REST: RESTful interface for programmatic access to PubChem. Nucleic Acids Res. 2018 July 2; 46(W1):W563-570. doi:10.1093/nar/gky294.
4. PUG VIEW. PubChem Docs [webiste]. Retrieved from https://pubchemdocs.ncbi.nlm.nih.gov/pug-view.
3. Kim S, Thiessen PA, Cheng T, Zhang J, Gindulyte A, Bolton EE. PUG-View: programmatic access to chemical annotations integrated in PubChem. J Cheminform. 2019 Aug 9; 11:56. doi:10.1186/s13321-019-0375-2.