---
title: "Querying PubChem with AnnotationGx"
author: "Christopher Eeles"
package: AnnotationGx
output: 
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Querying PubChem with AnnotationGx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

# Introduction to PubChem APIs

PubChem provides two publicly available RESTful APIs for programatically
accessing data in their databases. The first of these, the PubChem PUG REST API,
is designed to make specific queries based on some input identifier and return
data which PubChem has labelled or computed internally [1]. This API can be used 
to do things like get standardized PubChem identifiers CID or SID from 
drug names, external database identifiers, etc. The second is the PubChem PUG 
VIEW API [2]. This interface allows access to a variety of data which are 
aggregated by PubChem to display along with their own records, but are not 
curated by PubChem itself.

## PubChem PUG REST API

This package inlcudes functions for making arbitrary queries to the PubChem
REST API. The most fundamental of these is the `getRequestPubChem` function, 
which uses its parameters to construct a single GET request to the PUG REST API, 
returning an `httr::response` object. The documentation for this function
also includes the most detail about the structure of the PUG REST API, and
developers wishing to write custom queries for this package should review
the documentation with `?getRequestPubChem`.

For convenience, a wrapper function, `queryRequestPubChem`, is available which 
automatically converts the results to a `data.frame` using `jsonlite::fromJSON`. 
This function only works when `output` parameter is set to `JSON`. 

The major workhorse for PubChem queries in this package is the `queryPubChem` 
function. This function is able to batch queries such that multiple identifiers 
are mapped per GET request. It also implementes parallelized queries as well as
query throttling to ensure users do not exceed the 5 query per second, 400
query per minute limits imposed by PubChem. Both 
`queryRequestPubChem` use the `...` parameter to pass
arguments through to `getRequestPubChem`. As such, users should review the
`getRequestPubChem` documentation prior to using either query functions.

Two additional parameters are added in `queryPubChem`. The `batch` query 
tells the function whether to use multiple identifiers per GET request. This
parameter should be set to `TRUE` (the default) unless you are retrying failed 
queries. This is because batch queries fail completely if ANY of their
their ids fail to map. The `raw` parameter is included for debugging purposes,
This can be used to diagnose HTTP errors using the `headers(response)` function.

Use of `queryPubChem` is mainly intended for developer use, and beginners
with this package should focus on using the more specialized functions described
below. It is worth noting, however, that all specialized functions were 
constructed using `queryPubChem`.

The PUG REST API is divided into multiple domains, each representing a distinct
entity in the PubChem Schema. Details of these domains can be found in 
`?getRequestPubChem`. The three most important domains are discussed below.

### Querying the Compound Domain

The compound domain allows users to make queries on pharmacologically relevant
entities such as drugs. The main PubChem id for compounds is Compound IDentifier 
(CID). Each CID may map to one or more Substance IDentifiers (SID). In this
context, SID is a less stable identifier, possibly only contributed
by a single source and not yet standardized. CIDs on the otherhand, have stable
PubChem entries and normalized names which are generally supported accross 
multiple data sources. As such, newly discovered or rarely used chemicals
may have an SID but not a CID.

For most users, the compound domain will provide the bulk of the information
they need about their drugs of interest. The `getPubChemCompound` function
has been included to allow users to easily query this domain.

#### Common Use Cases

1. Get CID from Drug Names

```{r eval=FALSE}
library(AnnotationGx)
library(data.table)
drugs <- c("Erlotinib", "Acadesine", "Camptothecin", "Vincaleukoblastine", 
    "Cisplatin")
namesToCID <- getPubChemCompound(drugs, from='name', to='cids')
```

It is possible for names to multimap to CIDs. This is the case for 
'Vincaleukoblastine' in the above query. In cases of multimapping, usually
the first entry has the highest similarity to the requested drug. To subset 
to only the first occurrence of each of drug name:

```{r eval=FALSE}
# Using duplicated function
noDuplicatesNamesToCID <- namesToCID[!duplicated(name), ]
# Using group-by operations
noDuplicatesNamesToCID <- namesToCID[, first(.SD), by=name]
```

In cases with multimapping, we highly recommend you manually check that the
CID you select is optimal. This can be done with a simple look up in PubChem
via your web browser if there are only few multimapped drugs. You can also
keep all duplicates, then decide on the best option when fetching compound
properties like this:

```{r eval=FALSE}
multiIDs <- namesToCID[name == 'Vincaleukoblastine', ]$cids
CIDtoTitle <- getPubChemCompound(multiIDs, from='cid', to='property', 
    properties='Title')
```

Viewing the results, we can see that indeed the first CID is the best,
as the Title of the PubChem entry is exactly our original query. Multimapping
occurs because PubChem look-ups by name also consider synonyms.

For this reason, we recommend mapping from names to CID or SID first, then
using those official PubChem identifiers to fetch additional information about 
your compound of interest.

```{r eval=FALSE}
# Extract the mutlimapped IDs
multiIDs <- namesToCID[name == 'Vincaleukoblastine', ]$cids
```

2. Get SID from Drug Names
```{r eval=FALSE}
namesToSID <- getPubChemCompound(drugs, from='name', to='sids')
```

Multimapping is an even greater concern when fetching SIDs, since a single CID 
can map to many SIDs. As a result, we only recommend using SIDs for 
compounds which have no available CID.

3. Get Synonyms from Drug Names
```{r eval=FALSE}
namesToSynonyms <- getPubChemCompound(drugs, from='name', to='synonyms')
```

4. Get Compound Properties from CIDs
```{r eval=FALSE}
cids <- namesToCID$cids
propertiesFromCID <- getPubChemCompound(cids, 
    properties=c('Title', 'InChIKey', 'CanonicalSMILES'))
```

A short-list of properties is available in the function documentation, see
`?getPubChemCompound`. A link in the DETAILS section will take you to a full
list of available compound properties in the PubChem API documentation.

### Querying the Substance Domain

The substance domain allows users to query a broader range of chemicals than
the compounds domain. This can be useful if mappings from name or CID failed
to get the information you are interested in.

### Querying the Assay Domain

A functions for making specific queries to the assay domain on the PUG REST API
has not been implemented yet. A user could implement their own using the
`queryPubChem` functions and knowledge of the PubChem API. It is possible
we will add additional methods to query this domain to retrieve information
such as drug targets or doseresponse metrics.

### Specialized functions

To make the package more user friendly, we have included a number of convenience
functions which further limit the scope of queries, but require less knowledge
of the underlying structure of the PubChem API. We hope that this will allow
users with no knowledge or interest in details of the PubChem API to retrieve 
useful information nonetheless.

1. Mapping from NSC Number to CIDs

```{r eval=FALSE}
NSC_numbers <- c(1026L, 1027L, 1042L, 1059L, 1063L, 1139L)
NSCtoCID <- getPubChemFromNSC(NSC_numbers, to='cids')
```

Additional functions mapping from external identifiers will be added as the 
need arises. Additional documentation will be added to this section at that 
time.

### Handling Failed queries

Queries to the PubChem API can fail for various reasons. Maybe there is a typo
in your query identifiers, maybe some of the identifiers don't map to any
PubChem enteries, maybe there was a connection interuption during your query.

The functions in this package should provide a warning whenever queries fail.
The identifiers for failed queries will be available as an attribute on the
returned results.

## PubChem PUG VIEW API

The PUG View API is primarily designed to provide summary reports of all 
external identifiers for a given PubChem entry. Each summary provides the
values of external annotations, as well as links to their source database.

### Querying Individual Records

A function for querying individual PUG VIEW records has not been implemented yet.
We will update this documentation as soon as we have the feature avaiable.

### Querying Annotations

The PubChem PUG View Annotations API provides summaries of third-party 
annotations across all PubChem records. These annotations are linked To
PubChem compounds via CID and/or SID. This API does not allow specific
input queries, instead returning the entire record for the user to parse
and extract relevant information.

To determine what kind of annotations are available, you can get a list
of headers by calling the function with no arguments:

```{r eval=FALSE}
headers <- getPubChemAnnotations()
# %like% is a convenient way to do regex subsets on a data.table
#>it is equivalent to grepl('ATC', Heading)
headers[Heading %like% 'ATC', ]
```

You can see additional documentation for this function using 
`?getPubChemAnnotations` from the R console.

1. Getting All ATC Codes

```{r eval=FALSE}
CIDtoATC <- getPubChemAnnotations('ATC Codes')
```

2. Getting All NSC Numbers

```{r eval=FALSE}
CIDtoNSC <- getPubChemAnnotations('NSC Number')
```

3. Getting All CTD Codes

```{r eval=FALSE}
CIDtoCTD <- getPubChemAnnotations('CTD Chemical-Gene Interactions')
```

4. Getting All DILI Statuses

```{r eval=FALSE}
CIDtoDILI <- getPubChemAnnotations('Drug Induced Liver Injury')
```

5. Getting All CAS Registry Numbers

The PubChem PUG VIEW API has many records with valid CAS identifiers, therefore
this function can be very slow. Since it returns values for ALL PubChem records,
it should only need to be run once. Subsequent look-ups can be done from the
resulting `data.table` based on CID or SID. Make sure to save the file
as soon as this function returns to prevent having to run it again.

```{r eval=FALSE}
CIDtoCAS <- getPubChemAnnotations('CAS')
```

6. Getting All Compound Synonyms

While it is generally faster and more efficient to use the `getPubChemCompound`
function to look up synonyms based on some known identifier, it may be useful
to get an exaustive list of synonyms when other mapping methods fail.

```{r eval=FALSE}
headers[Heading %like% 'Syn']
```

Here we can see that there are two options for looking up compound synonyms
available via the PUG VIEW API. We can get results for both.

```{r eval=FALSE}
CIDtoNamesAndSyn <- getPubChemAnnotations('Names and Synonyms')
```

```{r eval=FALSE}
CIDtoSynAndIDs <- getPubChemAnnotations('Synonyms and Identifiers')
```

7. Other Annotations

The `getPubChemAnnotations` function supports arbitrary annotation headers.
We have included default parsing functions to take the raw query results
for the use cases above and return a well formatted table of result data.

However, users wishing to retrieve other annotations can do so. By defaults
the results will be returned as single `data.table` which the user will need
to process into something usable. If you find you are often retreiving the
same kind of annotations, it is also possible for developers to write a custom
parser function which will automatically format the results before they are
returned.

This function should be passed to the `parseFUN` parameter, and should accept
the unparsed return of the function and process it to the desired output. If
you have some efficient parse functions you would like to see added as
defaults in the package please feel free to open an Issue or Pull Request
on our GitHub page.

# References
1. PUG REST. PubChem Docs [website]. Retrieved from https://pubchemdocs.ncbi.nlm.nih.gov/pug-rest.
2. Kim S, Thiessen PA, Cheng T, Yu B, Bolton EE. An update on PUG-REST: RESTful interface for programmatic access to PubChem. Nucleic Acids Res. 2018 July 2; 46(W1):W563-570. doi:10.1093/nar/gky294.
4. PUG VIEW. PubChem Docs [webiste]. Retrieved from https://pubchemdocs.ncbi.nlm.nih.gov/pug-view.
3. Kim S, Thiessen PA, Cheng T, Zhang J, Gindulyte A, Bolton EE. PUG-View: programmatic access to chemical annotations integrated in PubChem. J Cheminform. 2019 Aug 9; 11:56. doi:10.1186/s13321-019-0375-2.