---
title: "Querying PubChem with AnnotationGx"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Querying PubChem with AnnotationGx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction to PubChem APIs

PubChem is a database of chemical molecules and their biological activities. 
It is a part of the National Center for Biotechnology Information (NCBI), 
which is a part of the National Institutes of Health (NIH). PubChem provides
a set of APIs to query its database. The `AnnotationGx` package provides a set 
of functions to query PubChem using these APIs.

The first of these APIs is the `PubChem PUG REST API` which is designed to 
- make specific queries based on some input *identifier* and return data which
PubChem has labelled or computed internally [1]. 
  - This API is useful for querying information about a specific chemical compound
such as getting the standardized PubChem identifier (CID) for a given chemical name or
smiles string, or getting the chemical structure for a given CID.
  - It provides access to a wide range of data including chemical properties, 
bioassay data, and chemical classification data, given a specific identifier.

The second API is the `PubChem PUG VIEW API` which is designed to:
- give accesse to aggregated annotations for a given chemical compound [3] that 
is mapped to their data, but not curated by PubChem itself. 
  - i.e it provides access to annotations from external sources such as
UniProt, ChEBI, and ChEMBL, given a specific identifier.


# Setup

```{r setup}
library(AnnotationGx)



```


# 

### Mapping from chemical name to PubChem CID
The main function that is provided by the package is `mapCompound2CID`. 

``` {r map aspirin to cid}
mapCompound2CID("aspirin")
```

In the case of a compound that can't be mapped, `NA` will be returned and a warning will be issued.

``` {r map non existent compound to cid}

(result <- mapCompound2CID(c("non existent compound", "another bad compound")))

failed <- attributes(result)$failed

# get the list of failed inputs 
names(failed)

# get the error message for the failed input
failed[1]

```

### Mapping from PubChem CID to Properties
TODO::


### Mapping from PubChem CID to Annotations
TODO::


# References
1. PUG REST. PubChem Docs [website]. Retrieved from https://pubchemdocs.ncbi.nlm.nih.gov/pug-rest.
2. Kim S, Thiessen PA, Cheng T, Yu B, Bolton EE. An update on PUG-REST: RESTful interface for programmatic access to PubChem. Nucleic Acids Res. 2018 July 2; 46(W1):W563-570. doi:10.1093/nar/gky294.
4. PUG VIEW. PubChem Docs [webiste]. Retrieved from https://pubchemdocs.ncbi.nlm.nih.gov/pug-view.
3. Kim S, Thiessen PA, Cheng T, Zhang J, Gindulyte A, Bolton EE. PUG-View: programmatic access to chemical annotations integrated in PubChem. J Cheminform. 2019 Aug 9; 11:56. doi:10.1186/s13321-019-0375-2.