---
title: "Querying Unichem Database"
author: 
  - name: Jermiah Joseph, Shahzada Muhammad Shameel Farooq, and Christopher Eeles
output: rmarkdown::html_vignette 
vignette: >
  %\VignetteIndexEntry{Querying Unichem Database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL ## Related to https://stat.ethz.ch/pipermail/bioc-devel/2020-April/016656.html
)
```


# Introduction to the Unichem API

The UniChem database provides a publicly available REST API for
programmatic retrieval of mappings from standardized structural compound
identifiers to unique compound IDs across a range of large online
cheminformatic databases such as PubChem, ChEMBL, DrugBank and many more.
The service accepts POST requests to two different end-points:
`/compound` and `/connectivity`. Both endpoints accept query parameters via the
POST body in JSON format. The `/compound` API returns exact matches for the
queried compound, while the `/connectivity` API uses layers of the International
Chemical Identifier (InChI) of the query compound to return exact matches as
well as structurally related compounds such as isomers, salts, ionizations
and more.
[@UniChemBeta; @chambersUniChemUnifiedChemical2013]

The functions in `AnnotationGx` have been designed to allow package users to
easily query UniChem resources without any pre-existing knowledge of
HTTP requests or the API specifications. In doing so we hope to provide an
R native interface for mapping between various cheminformatic databases,
accessible to anyone familar with using R functions!

```{r load_pkg_example}
library(AnnotationGx)
```


# Available Databases

To see a table of database identifiers available via UniChem, you can call
the `getUniChemSources` function. 
By default, just the database shortname ("Name") and UniChem's ID for it ("SourceID") columns 
are returned.
To return all columns, pass the `all_columns = TRUE` argument 

```{r get_sources_short_echo}
getUnichemSources()
```

When mapping using the `queryUnichemCompound` function, these are the sources that can be used from, 
and the databases to which the compound mappings will be returned. 

# Querying UniChem Compound API

The `queryUnichemCompound` function allows you to query the UniChem Compound API
to retrieve mappings for a given compound identifier. The function takes two mandatory arguments.
The first is the `compound` argument which is the compound identifier to be queried. 
The second is the `type` argument which is the type of compound identifier to search for.
Options are "uci", "inchi", "inchikey", and "sourceID". 
The `sourceID` argument is optional and is only required if the `type` argument is "sourceID".

The function returns a list of:

1. "External_Mappings" `data.table` containing the mapping to other Databases with the following headings:
   1. "compoundID" `character` The compound identifier
   2. "Name" `character` The name of the database
   3. "NameLong" `character` The long name of the database
   4. "SourceID" `character` The UniChem Source ID
   5. "sourceURL" `character` The URL of the source
2. "UniChem_Mappings" `list` of the following six mappings:
   1. "UCI" `character` The UniChem Identifier
   2. "InchiKey" `character` The InChIKey
   3. "Inchi" `character` The InChI
   4. "formula" `character` The molecular formula
   5. "connections" `character` connection representation "1-6(10)13-8-5-3-2-4-7(8)9(11)12"
   6. "hAtoms" `character`  hydrogen atom connections "2-5H,1H3,(H,11,12)"


#### Example Searching using `uci` (UniChem Identifier) 
Note: This type of query requires you to know the UniChem Identifier for the compound. 

```{r uci query}

queryUnichemCompound(compound = "161671", type = "uci")

```


