---
title: "Annotating CTRP Treatments"
---

# Introduction

This vignette compares annotating CTRP-provided treatment ids to PubChem CIDs and CTD information.

Whereas the PubChem CID is a unique identifier for a compound, the PubChem API does not easily map 
treatment names to CIDs, atleast not in a way that is easily scalable.

The [CTD2](https://ctd2-dashboard.nci.nih.gov/dashboard) database is the central database where CTRP 
data is hosted. They happen to expose (an API)[https://ctd2-dashboard.nci.nih.gov/dashboard/#api-documentation] 
for their database. The API calls they describe on their API documentation is useful, but they have an endpoint:
`GET /compound/{compoundId}` that is not documented. This endpoint is useful for mapping compound names to CIDs.

The functionality for this is implemented in the `mapCompound2CTD` function.

It is an investigation to see which of the methods might map more compounds

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(AnnotationGx)

data(CTRP_treatmentMetadata)
```



``` {r test_both}

# get a random row from the CTRP_treatmentMetadata

treatment <- CTRP_treatmentMetadata[sample(1:nrow(CTRP_treatmentMetadata), 1),CTRP.treatmentid]
sprintf("CTRP treatment id : %s", treatment)

# map the treatment to a CID using the CTD database
mapCompound2CTD(treatment)[, .(displayName, PUBCHEM)]


# map the treatment to a CID using PubChem
mapCompound2CID(treatment)

```


# Annotating using the CTD database


```{r run_CTD}
result <- CTRP_treatmentMetadata[, mapCompound2CTD(CTRP.treatmentid, query_only = F, raw = F)]

show(result)
message("Failed results: ", result[is.na(result$PUBCHEM), .N])

failed_names <- result[is.na(result$PUBCHEM),displayName]
```


# Annotating using PubChem
``` {r run_CTRP_Pubchem, eval = FALSE}
(compounds_to_cids <- 
  CTRP_treatmentMetadata[, 
    AnnotationGx::mapCompound2CID(
        names =  CTRP.treatmentid,
        first = TRUE
        )
      ]
)
failed <- 
  attributes(compounds_to_cids)$failed |> 
    names()
```


``` {r Pubchem Failed, eval = FALSE}
failed <- unique(CTRP_treatmentMetadata[CTRP.treatmentid %in% failed, ])

failed[, CTRP.treatmentid_CLEANED := cleanCharacterStrings(CTRP.treatmentid)]

(failed_to_cids <-
  failed[, 
    AnnotationGx::mapCompound2CID(
      names = CTRP.treatmentid_CLEANED,
      first = TRUE
    )
  ]
)
failed_again <-
  attributes(failed_to_cids)$failed |> 
    names()

```

``` {r pubchemfailed again, eval = FALSE}
failed_dt <- merge(failed_to_cids[!is.na(cids),], failed, by.x = "name", by.y = "CTRP.treatmentid_CLEANED", all.x = F)
failed_dt$name <- NULL

successful_dt <- merge(CTRP_treatmentMetadata, compounds_to_cids[!is.na(cids),],by.x = "CTRP.treatmentid", by.y = "name",  all.x = F)

mapped_PubChem <- data.table::rbindlist(list(successful_dt, failed_dt), use.names = T, fill = T)

```
